DOpElib: A Goal Oriented Software Library for Computing PDEs and Optimization Problems

### 2012-11-07 Paper_May_2012/main.tex
Some comments to the new sections added.

\subsection{Solving of optimization problems}\label{sec:opt}
Upgrading the framework for solving optimization problems
is described in this section. A prototypical problem reads:
\begin{align*}
\min\;&J(q,u) \\
  &\text{s.t.}\; a(q,u)(\phi) = 0 \quad \forall \phi\in V,\\
  &a \le q \le b,\\
  &g(q,u) \le 0,  
\end{align*}
where $u$ is a FE-function and $q$ can either be a FE-function or some 
fixed number of parameters, $a$ and $b$ are constraint bounds for the control $q$,
and $g(\cdot)$ is some state constraint.
To keep the presentation brief we will neglect the additional bounds
\[
a \le q \le b,\qquad g(q,u) \le 0
\]
although they can be treated by \dope as well.
Then, in order to minimize the functional $J$ one may apply the, so-called, 
reduced approach, in which we eliminate the dependent variable $u$ by the solution 
operator $S$ of the PDE, i.e., $a(q,S(q))(\phi) = 0 \quad \forall \phi\in V$. 
This leads to the problem
\begin{align*}
\min\;&j(q) = J(q,S(q)). 
\end{align*}
Then, one can apply any algorithm to solve the, unconstraint, minimization 
problem. This, will in any case require the evaluation of $j(q)$ which 
we can do, e.g., using \texttt{StatPDEProblem::ComputeReducedFunctional} for 
a stationary PDE. If one would like to use derivative based methods, one 
needs in adition derivatives of $j$ and thus of $J$ as well as of $S$. 
To do so, we remark, that one can calculate the derivative of 
$j$ using the adjoint calculus. Then the directional derivative 
of $j$ in direction $\delta q$ is given by
\[
j'(q)(\delta q) = J'_q(q,S(q))(\delta q) + (S^* J_u'(q,S(q)),\delta q)
\]
where the index $q$ or $u$ denote directional derivatieves for the 
respective argument. To calculate action of the adjoint operator $S^*$ 
we note that $z =  S^* J_u'(q,S(q))$
can be computed via the adjoint PDE
\begin{equation}
  \label{opt_KKT_system}
  \begin{aligned}
    a_{u}'(q,u)(\phi , z) &= J'_{u}(q,u)(\phi) & \forall \phi&\in {\cal X}.
  \end{aligned}
\end{equation}
Similarly formulas for the second derivatives can be obtained.

>>>>

\subsection{Solving of optimization problems}\label{sec:opt}
Upgrading the framework for solving optimization problems
is described in this section. A prototypical problem reads:
\begin{align*}
\min\;&J(q,u) \\
  &\text{s.t.}\; a(q,u)(\phi) = 0 \quad \forall \phi\in V,\\
  &a \le q \le b,\\
  &g(q,u) \le 0,  
\end{align*}
where $u$ is a FE-function and $q$ can either be a FE-function or some 
fixed number of parameters, $a$ and $b$ are constraint bounds for the control $q$,
and $g(\cdot)$ is some state constraint.
To keep the presentation brief we will neglect the additional bounds
\[
a \le q \le b,\qquad g(q,u) \le 0
\]
although they can be treated by \dope{} as well.
Then, in order to minimize the functional $J$ one may apply the, so-called, 
reduced approach, in which we eliminate the dependent variable $u$ by the solution 
operator $S$ of the PDE, i.e., 
\begin{align*}
a(q,S(q))(\phi) = 0 \quad \forall \phi\in V. 
\end{align*}
This leads to the problem
\begin{align*}
\min\;&j(q) = J(q,S(q)). 
\end{align*}
Then, one can apply any algorithm to solve the, unconstraint, minimization 
problem. This, will in any case require the evaluation of $j(q)$ which 
we can do, e.g., using \texttt{StatPDEProblem::ComputeReducedFunctional} for 
a stationary PDE. If one would like to use derivative based methods, one 
needs in adition derivatives of $j$ and thus of $J$ as well as of $S$. 
To do so, we remark, that one can calculate the derivative of 
$j$ using the adjoint calculus. Then the directional derivative 
of $j$ in direction $\delta q$ is given by
\[
j'(q)(\delta q) = J'_q(q,S(q))(\delta q) + (S^* J_u'(q,S(q)),\delta q)
\]
where the index $q$ or $u$ denote directional derivatieves for the 
respective argument. To calculate action of the adjoint operator $S^*$ 
we note that $z =  S^* J_u'(q,S(q))$
can be computed via the adjoint PDE
\begin{equation}
  \label{opt_KKT_system}
  \begin{aligned}
    a_{u}'(q,u)(\phi , z) &= J'_{u}(q,u)(\phi) & \forall \phi&\in {\cal X}.
  \end{aligned}
\end{equation}
Similarly formulas for the second derivatives can be obtained.
